---
title: "MyTitle"
author: "NickM"
date: "4 May 2017"
output: 
  ioslides_presentation:
  smaller: true
  transition: faster
  fig_width: 8
  fig_height: 6
---
<!-- <# output:   colon at aend -->
<!-- #   outputFormat: colon at end -->
<!-- #       outputformatoptions -->
<!-- http://rmarkdown.rstudio.com/ioslides_presentation_format.html#overview -->

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
```

## R Markdown

<div class="red2">
This is an R Markdown presentation. Markdown is a simple formatting syntax for authoring HTML, PDF, and MS Word documents. For more details on using R Markdown see <http://rmarkdown.rstudio.com>.
</div>

When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document.

## Slide with code
Preamble to code
```{r, comment="",}
head(mtcars)
mtcars$disp * 2
```
## Plotting in Knitr
```{r, echo = FALSE, fig.align= 'center', fig.cap='fish fishfish',background = "tan4"}
plot(mtcars$wt, mtcars$mpg, xlab="Weight", ylab="MPG")
```